The Markov Link Method: a nonparametric approach to combine observations from multiple experiments

### 2018-08-07 paper/paper.tex
Back to bayes

% \newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
% \newcommand{\figgygrs}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth]{#2}\\#3\end{tabular}}
% \newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
% \newcommand{\figgygrsH}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth,angle=270]{#2}\\#3\end{tabular}}

\newcommand{\figgygr}[2]{pic}
\newcommand{\figgygrs}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
\newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
\newcommand{\figgygrsH}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}

@@

If we have a single large experiment that measures each specimen with both measurement procedures, we can estimate the link without the MLM.  In this case it is straightforward to study the link between the two perspectives, because we can examine specimens through both perspectives simultaneously.  However, in many cases we have some experiments in which specimens are measured with one procedure, other experiments using the other procedure, and few or no experiments where specimens are measured with both procedures.  This is particularly true if the measurement techniques destroy or alter the specimen in the process of measuring it.

@@

Imagine that we have a large collection of human cells.  For each cell we are interested in 
%
\begin{enumerate}
    \item $\ell$, the sampling procedure used to obtain the specimen.  For example, we can use select different subpopulations of cells based on proteomic markers.  For each marker we might have a different sampling procedure designed to select cells with that marker.
    \item $X$, the cell's `transcriptomic type,' as measured by looking at the transcriptomic activity in the cell.  
    \item $Y$, the cell's `morphological type,' as measured by looking at an image of the cell.
\end{enumerate}
%
Our task is to reconcile the two different notions of `cell-type' indicated by $X$ and $Y$.  We would like to know what transcriptomic types are associated with what morphological types.  For example, we might like to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  Our main assumption is that these kinds of associations are unaffected by the choice of sampling procedure.  In particular:

@@

Here $\mathbb{P}(Y|X,\ell)$ indicates the probability that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.  The assumption is that this probability actually does not depend upon $\ell$.  This can be understood intuitively through a thought experiment.  Let us imagine we have attained perfect understanding of cellular biology.  Someone selects a cell using one of a set of sampling procedures and measures the transcriptomic activity of the cell.  We are then told the cell's transcriptomic activity (but we are not told the sampling procedure used to gather the specimen).  Using our perfect knowledge of the physical systems, we could then make predictions about a measurement of the cell's morphology.  Now we are told new information: we are told the sampling procedure used to gather the specimen.  Would this substantially change our predictions?  If so, the MLM assumption is violated.  However, if the transcriptomic measurement is sufficiently informative, learning the sampling strategy would not substantially change our predictions about the cell morphology.  As long as we can find a measurement procedure that is sufficiently informative in this way, we can apply the MLM.  A list of examples where the MLM might apply may be found in Section \ref{sec:examples}.

@@

In the specific case we will focus on, all of the experimental data can be summarized with two matrices, $N^X$ and $N^Y$.  Here $N^X_{\ell x}$ indicates the number of specimens gathered with procedure $\ell$ that have transcriptomic type $x$.  Likewise $N^Y_{\ell y}$ indicates the number of specimens gathered with procedure $\ell$ that have morphological type $y$.

Taking a frequentist point of view, the Markov Link Method uses this data to produce two kinds of objects:
\begin{itemize}
    \item $\tilde q(y|x)$, a point estimate for the link parameter $q^*(y|x)$.  To obtain this we must incorporate nuisance parameters: let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$.  The log likelihood of the data is then given by
    \[
    L(p,q)=\left(\sum_{\ell,x} N^X_{\ell x} \log p(x|\ell) + \sum_{\ell,y} N^Y_{\ell y} \log \left(\sum_x p(x|\ell) q(y|x)\right)\right)
    \]
    We use a maximum likelihood approach to build our point estimators:
    \[
    \tilde p,\tilde q \in \argmax_{p,q} L(p,q)
    \]
    We caution that this optimization problem does not always have a unique maximizer, so the result may depend upon the maximization strategy used.  Precise details for our optimization algorithms may be found in Appendix \ref{sec:optproblems}.  This non-uniqueness is related certain so-called `identifiability' issues: in some cases it may be impossible to obtain a perfect point estimate even with an infinite number of samples.   We discuss this issue at length in Appendix \ref{sec:casestudies}.  In light of these concerns, it is particularly important to also consder our uncertainty:
    %
    \item $C_{x,y}$, a 95\% confidence interval for $q^*(y|x)$.  We obtain these with profile likelihood approach.  The profile likelihood is defined as
    \[
    L_{x y}(\xi)=\max_{\substack{p,q \\ q(y|x)=\xi}} L(p,q)
    \]
    This is the maximum possible likelihood of the data under the constraint that $q(y| x)=\xi$.  We then take $L^* = \sup_{p,q} L(p,q)$ and define a confidence region for $q^*( y| x)$ by
    \[
    R_{x y} = \left\{\xi:\ L^* - L(\xi) < k\right\}
    \]
    We choose $k$ to achieve the desired coverage.  That is, we would like to choose $k$ so that $q^*(y|x) \in R_{x y}$ with probability at least 95\%.  We refer the reader to Appendix \ref{sec:confinterval} for details on choosing $k$.  Other coverages besides 95\% can also be produced by the same method.  Finally, for simplicity of visualization we transform the set $R_{x y}$ (which could be composed of several disjoint intervals) into a simple interval by taking $C_{x,y} = [\min R_{x,y},\max R_{x,y}]$.  Since this interval is a superset of the region, it will certainly contain $q^*(y|x)$ as long the region does.
\end{itemize}

A Bayesian point of view would suggest a different procedure, one which might provide a useful complementary perspective.  This is worthy of future research.  However, we caution that it is nontrivial to choose reasonable prior distributions for $p^*,q^*$.  Priors which may seem uninformative (such as the Dirichlet distribution) can have significant unanticipated consequences, due to the high-dimensional nature of the problem.  We give an example in \ref{sec:bayesian}.

The frequentist confidence interval allows us to avoid making any modeling assumptions (beyond the basic Markov Link Method assumption).  However, it does come with some drawbacks.   For example, we here take the view that values inside the confidence interval are somehow `plausible' or `likely.'  Such a statement cannot be philosophically justified without some at least a hint of Bayesian reasoning (cf.\ \cite{morey2016fallacy}).  It is also important to note that, as is typical for confidence intervals, we make no corrections for multiple testing.  We therefore expect that a small proportion of our confidence intervals will not contain the truth.   It is therefore important to consider all of the intervals produced and be wary of `cherry picking' only those intervals that support a particular conclusion.  However, if cherry picking is desired, a false discovery rate correction may be applied (cf.\ \cite{benjamini2005false}).  A related drawback is that the confidence intervals are not always logically consistent with one another.  For example, it can happen that that $C_{x,y_1}=[0,1]$ and $C_{x,y_2}=[.2,1.0]$.  These intervals don't really make sense together, in that if $q(y_2|x)\geq .2$ then certainly $q(y_2|x)\leq .8$.  A possible remedy would be to construct a simultaneous confidence region for all the parameters concerning a fixed value of $x$.  Analyzing the statistical properties of these high-dimensional confidence regions would be nontrivial; we leave it for future research.

@@

\caption{{\bf When can the MLM accurately estimate the link?} \label{fig:simI}  We use simulations to test the MLM.  In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, sampling procedures, and the link parameters.  In each trial our `ground truth' link parameters picked uniformly at random from the parameter space.  The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct sampling procedures.  In subfigure (b) we consider the MLM's 95\% confidence intervals.  For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM confidence interval $C_{x,y}$.  We show the interval (in black) centered around the ground truth (in red).  Since the link parameters lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).  The interval mostly contain the truth, and get narrower with more samples as long as there are enough distinct sampling procedures.   In subfigure (c) we focus on the case of exactly $4$ sampling procedures.  In this case it is only possible to get very accurate estimates if the link takes on a particular form.  If the link is indepenent (i.e.\ $X,Y$ are independent) or if the link is invertible and deterministic (e.g.\ $X=Y$) then we can determine what the link is.  However, if the truth is more complex, small numbers of sampling procedures may make it impossible to determine the true link; this issue is discussed in detail in Section \ref{sec:casestudies}.}
}

We used simulations to test the two parts of the MLM: the point estimators $\tilde q(y|x)$ and confidence intervals $C_{x,y}$.  For any fixed value of the ground truth parameters, $(p^*,q^*)$, we would ordinarily hope for two things:
\begin{description}
    \item[Estimator convergence] As we obtain more samples, $\tilde q(y|x)$ should converge to $q^*(y|x)$ for each $x,y$.
    \item[Conservative coverage] $\mathbb{P}(q^*(y|x) \in C_{x,y})\geq95\%$ for each $x,y$.  That is, for each choice of $x,y$ the confidence interval $C_{x,y}$ should contain $q^*(y|x)$ with probability at least 95\%.\footnote{Note that we are not producing a `simultaneous' interval.  In a simultaneous interval, there would be a 95\% chance that $q^*(y|x)\in C_{x,y}$ for \emph{every} $x,y$.  The key difference is the change from `each' (in our intervals) to `every' (in simultaneous intervals).  Thus, it may be that $q^*(y|x)\notin C_{x,y}$ for a few values of $x,y$.  For \emph{most} values of $x,y$ the true parameters should be contained inside of their corresponding confidence intervals.}
\end{description}
Our simulations yielded good news and bad news.  The good news: the MLM generally achieves conservative coverage.  In many cases, the MLM also achieves estimator convergence.  The bad news: in many cases the MLM fails to achieve estimator convergence.  Indeed, we can show that an `identifiability issue' blocks any possibility of success in some cases; we detail this problem in Appendix \ref{sec:casestudies}.  The following two simulations give a good overview of the relevant issues:

\begin{itemize}
    \item How does MLM performance depend on the number of samples and the number of sampling procedures?  We suppose we have two measurement procedures which return categorical measurements among six categories.  Thus the first procedure yields $X\in \{1,2,3,4,5,6\}$ and the second yields $Y\in \{1,2,3,4,5,6\}$.  To see how the method performs in different circumstances, we will run many trials.  In each trial we fix the number of sampling procedures and pick `ground truth' parameters uniformly at random from the parameter space.  We then simulate a dataset from these ground truth parameters, fixing the total number of samples (spreading these samples equally among all combinations of sampling procedure and measurement procedure).  Finally, we apply the MLM to the simulated data to get the point estimates $\tilde q(y|x)$ and the confidence intervals $C_{x y}$.  We measure the overall estimator convergence using a kind of total variation distance:
    \[
    \mathrm{Error}(\tilde q,q^*)=\frac{1}{12}\sum_{x=1}^{6}\sum_{y=1}^{6} |\tilde q(y|x) - q^*(y|x)|
    \]
    This error ranges between zero and one.  Zero error indicates that $\tilde q=q^*$. An error of one indicates that the estimate has completely incorrect beliefs about the probability mass, i.e. $q^*(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.  We also look at the MLM confidence intervals and see how often they cover the true parameters.

    The results are summarized in Figure \ref{fig:simI}.  In trials with more samples, the estimator generally has lower error.  However, to make the error actually converge to zero we need at least six distinct sampling procedures.  This figure also shows that the confidence intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.  A more precise analysis of these intervals suggests that they achieve conservative coverage (see Appendix \ref{sec:confinterval} for details).  With many samples and sampling procedures, these intervals are small and concentrated around the truth.  With fewer samples or sampling procedures, the intervals are sometimes forced to be larger.  However, our uncertainty is not the same for every parameter.  In some cases we are able to obtain a very tight confidence interval for one particular parameter even though the overall estimator error is high.

    \item How does estimator convergence depends upon the link parameters themselves?  In each trial of this simulation we use four sampling procedures and our measurement procedures yield one of six categories.  In the previous simulations we saw that estimator convergence was generally impossible in this situation, due to the small number of sampling procedures.  However, in our previous simulations we picked the link parameters uniformly from parameter space.  Now we will be more choosey.   On one extreme, we will produce trials where the link makes $X,Y$ independent, i.e. $q^*(y|x)=q^*(y|x')$ for every $x,x'$.  On the other extreme, we will have trials where $X,Y$ are deterministically related by the equation $X=Y$.  We will also consider every link ``in-between'' these two extremes (found by convex combinations).  Figure \ref{fig:simI} shows that estimator convergence is actually possible in the two extreme cases.  However, estimator convergence fails for the in-between cases.  We leave a complete mathematical understanding of this result to future work.  For the purposes of this paper, we content ourselves by noting one interesting case: let the measurements return one of $2^k$ categories and let us use only $k+1$ carefully chosen sampling procedures.  Now suppose the link defines any invertible deterministic function between $X$ and $Y$.  In this case, with enough enough samples, we can determine both that the relationship is deterministic and the exact specification of the invertible function.  This result is proven in Appendix \ref{sec:casestudies}, Theorem \ref{thm:miracle}.

@@

\figgy{
\hfill{}\figgygr{.95}{../images/allenNlxy}\hfill{}
\caption{{\bf The input to the Markov Link Method: one experiment for each measurement procedure and each sampling procedure.}  Here we show a portion of the real-world data to which we applied the MLM.  The Allen Institute gathered neurons from the visual cortex of mice, using a variety of cre/lox-based sampling procedure (cf.\ \cite{tasic2017shared}).  Each strategy was designed to bring out different subpopulations of cells.  Neurons were measured to determine their `type,' using one of two procedures: Standard or Patch.  Standard recognizes 104 different types of neurons.  Patch has a coarser notion of cell-type, distinguishing only 10 types.  For each experiment, we tabulated the number of cells assigned to each type.  Above we show a subset of these results through heatmaps; the color of each square indicates the number of specimens found to have a particular type.  Using this kind of data, our task is to calibrate the two classification protocols.  That is, we want to be able to ask question of the following form: ``if a neuron is classified as being of type `Peri Kcnj8' by tool I, how might it have been classified by tool II?'' \label{fig:alleninput}}
}

@@

Our motivation for this problem arose from the Allen Institute's investigation of neuronal cell-types.  The institute has a variety of methods for examining a neuron and determining what kind of neuron it is.  However, many of these procedures destroy the neurons in the process of measuring them.  We'll look at two procedures in particular, which we'll call `Standard' and 'Patch.'  Standard uses a single-cell RNA sequencing pipeline to determine transcriptomic expression in the cell, and determine the cell-type based on these results.  Patch also uses transcriptomic information from RNA sequencing, but Patch additionally obtains electrophysiological and morphological properties of the neuron.  This comes at the cost of a degraded transcriptomic signal, requiring new methods to estimate the cell-type.  We refer the reader to \citep{tasic2017shared} for details on the two methods.

@@

Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: a variety of sampling procedures.  Using a cre/lox system (cf.\ \citep{tasic2017shared}) they were able to pick out various overlapping subpopulations of neurons using different procedures.  Each procedure was expected to yield different proportions of the different cell-types.  For each sampling procedure and eaach measurement procedure, many specimens were sampled and their cell-types determined.  The result of this process was two tables, parts of which are shown in Figure \ref{fig:alleninput}.  While it seemed clear that these tables should say something about the calibration, it was not obvious how to best use this information.  It was for this purpose that the MLM was developed.

@@

In future work, it would be interesting to use additional assumptions to help us estimate calibrations.  For example, as it stands the Markov Link Method will only yield narrow confidence intervals if each measurement tool returns one of a finite number of measurement outcomes.  In some cases, we may believe that similar measurement values should have similar probabilities.  Such smoothness assumptions would make it possible to apply the MLM to measurement tools which can return many different values.  The frequentist point of view allows us to phrase these kinds of assumptions as hard constraints that we believe in, instead of guesses about what might be more or less likely.  In some cases these assumptions are easier to think about than prior distributions, as we see in Appendix \ref{sec:bayesian}.

More generally, it is clear that good assumptions can help us get real insight for tough problems. Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, confidence regions can be constructed.  By probing these regions carefully, we can learn what the data actually has to say and what experiments will help us learn more.

@@

\section{Confidence interval details}

\label{sec:confinterval}

In the main text of this paper, Section \ref{sec:mlm}, we outline a method for producing confidence intervals for the link parameters $q^*(\tilde y|\tilde x)$ for a specific choice of $\tilde x,\tilde y$.  Here we give full details on this method and discuss its properties in depth.

>>>>

\newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
\newcommand{\figgygrs}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth]{#2}\\#3\end{tabular}}
\newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
\newcommand{\figgygrsH}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth,angle=270]{#2}\\#3\end{tabular}}

% \newcommand{\figgygr}[2]{pic}
% \newcommand{\figgygrs}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
% \newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
% \newcommand{\figgygrsH}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}

@@

If we have a single large experiment that measures each specimen with both measurement procedures, it is straightforward to estimate the link.   We can examine specimens through both perspectives simultaneously.  However, in many cases we have some experiments in which specimens are measured with one procedure, other experiments using the other procedure, and few or no experiments where specimens are measured with both procedures.  This is particularly true if the measurement techniques destroy or alter the specimen in the process of measuring it.

@@

Imagine that we have a large collection of human cells.  For each cell we are interested in 
%
\begin{enumerate}
    \item $\ell$, the sampling procedure used to obtain the specimen.  For example, we can select different subpopulations of cells based on proteomic markers.  For each marker we might have a different sampling procedure designed to select cells with that marker.
    \item $X$, the cell's `transcriptomic type,' as measured by looking at the transcriptomic activity in the cell.  
    \item $Y$, the cell's `morphological type,' as measured by looking at an image of the cell.
\end{enumerate}
%
Our task is to reconcile the two different notions of `cell-type' indicated by $X$ and $Y$.  We would like to know what transcriptomic types are associated with what morphological types.  For example, we might like to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  Our main assumption is that these kinds of associations are unaffected by the choice of sampling procedure.  In particular:

@@

Here $\mathbb{P}(Y|X,\ell)$ indicates the probability that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.  The assumption is that this probability actually does not depend upon $\ell$.  This can be understood intuitively through a thought experiment.  Let us imagine we have attained perfect understanding of cellular biology.  Someone selects a cell using one of a set of sampling procedures and measures the transcriptomic activity of the cell.  We are then told the cell's transcriptomic activity (but we are not told which of the sampling procedured was used to gather the specimen).  Using our perfect knowledge of the physical systems, we could then make predictions about a measurement of the cell's morphology.  Now we are told new information: we are told which sampling procedure was used to gather the specimen.  Would this substantially change our predictions?  If so, the MLM assumption is violated.  However, if the transcriptomic measurement is sufficiently informative, learning the sampling strategy would not substantially change our predictions about the cell morphology.  As long as we can find a measurement procedure that is sufficiently informative in this way, we can apply the MLM.  A list of examples where the MLM might apply may be found in Section \ref{sec:examples}.

@@

In the specific case we will focus on, all of the experimental data can be summarized with two matrices, $N^X$ and $N^Y$.  Here $N^X_{\ell x}$ indicates the number of specimens gathered with procedure $\ell$ that have transcriptomic type $x$.  Likewise $N^Y_{\ell y}$ indicates the number of specimens gathered with procedure $\ell$ that have morphological type $y$.  For each transcriptomic type $x$ and morphological type $y$, the Markov Link Method uses the data to produce two objects:
%
\begin{enumerate}
    \item $\hat q(y|x)$ -- a point estimate for the link parameter $q^*(y|x)$
    \item $C_{x,y}$ -- a credible interval which contains the true link parameter $q^*(y|x)$ with high probability
\end{enumerate}
%
To produce these objects we proceeds as follows.  Let $\tilde P(\cdot |\ell) \sim \mathrm{Dirichlet}(N^X_\ell+1)$ and $\tilde H(\cdot |\ell) \sim \mathrm{Dirichlet}(N^Y_\ell+1)$ for each $\ell$.  For any pair $h,h'$ let 
\[
D(h|h') = \sum_{\ell,y} h(y|\ell) \log \frac{h(y|\ell)}{h'(y|\ell)}
\]
denote the Kullback-Leibler divergence from $h$ to $h'$.  For each $\tilde P,\tilde H$, let $E(\tilde P,\tilde H)$ denote the minimum value of $D(\tilde H|h')$ among all $h'$ such that there exists $q$ satisfying $h'(y|\ell)=\sum_x \tilde P(y|\ell)q(y|x)$.  Let $P,H$ denote variables carrying the distribution of $\tilde P,\tilde H$ after conditioned on $E(\tilde P,\tilde H)=0$.   Let $\Theta(p,h)=\{q:\ h(y|\ell) = \sum_x p(x|\ell)q(y|x)\}$.  We then define the MLM objects as follows:
\begin{enumerate}
    \item $\hat q\triangleq\mathbb{E}[\argmin_{q\in \Theta(P,H)} \sum_{x,y} q(y|x)^2]$.
    \item The lower bound of $C_{x,y}$ is defined as the largest number $c$ such that $\min_{q\in \Theta(P,H)} q(y|x) \geq c$ with probability 97.5\%.  Likewise the upper bound is the smallest number $c$ such that $\max_{q\in \Theta(P,H)} q(y|x) \leq c$ with probability 97.5\%.
\end{enumerate} 
%
Code to compute these objects is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.

Our approach arises naturally from a Bayesian point of view.  To understand them, first notice that there are several `nuisance parameters' we must consider in order to estimate the link $q^*$.  Let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$ and $h^*(y|\ell) \triangleq \sum_x p^*(x|\ell)q(y|x)$ denote the probability such a cell will have morphological type $y$.  Note that the distribution of the data $N^X,N^Y$ is entirely determined by $p^*,h^*$.   In particular, if someone changed $q^*$ without changing $p^*,h^*$ there would be no way for us to know.  If this can occur, $q^*$ is said to be `nonidentifiable.'   Therefore, instead of trying to directly estimate $q^*$ (which may be impossible to accurately estimate) we focus on three quantities which can be estimated from data:
%
\begin{align*}
q^*_{\mathrm{mid}} &\triangleq \argmin_{q\in \Theta(p^*,h^*)} \sum_{x,y} q(y|x)^2\\
q^*_{\mathrm{lo},x,y} &\triangleq \min_{q\in \Theta(p^*,h^*)} q(y|x) \\
q^*_{\mathrm{hi},x,y} &\triangleq \max_{q\in \Theta(p^*,h^*)} q(y|x) 
\end{align*}
%
Note that all three of these optimization problems have unique solutions and would be straightforward to solve if $p^*,h^*$ were known; optimization details can be found in Appendix \ref{sec:optproblems}.  Moreover, regardless of any identifiability issues for $q^*$ itself, the objects $q^*_{\mathrm{lo},x,y}$ and $q^*_{\mathrm{hi},x,y}$ are always identifiable for every $(x,y)$ and define bounds on the true value of $q^*(y|x)$.

The MLM algorithm therefore adopts a Bayesian point of view to estimate $q^*_{\mathrm{mid}}$ and create a credible interval for $q^*_{\mathrm{lo},x,y},q^*_{\mathrm{hi},x,y}$.  We begin with uniform prior beliefs for the parameters of $p^*,h^*$.  That is, our prior is that $p^*(\cdot|\ell),h^*(\cdot|\ell)$ carry $\mathrm{Dirichlet}(1)$ distributions for each $\ell$.  Following the Bayesian philosophy, we would like to incorporate any additional knowledge through conditioning.  We have two important pieces of knowledge we would like to incorporate: that the MLM assumption holds (i.e.\ there is some $q^*$ so that $\sum_x p^*(x|\ell)q^*(y|x)=h^*(y|\ell)$) and we have observed the data $N^X,N^Y$.  By conditioning on this knowledge, we obtain the posterior distribution carried by the variables $P,H$ described above.  In terms of these variables, the MLM estimator $\hat q$ is simply the posterior mean for $q^*_{\mathrm{mid}}$ and the MLM credible interval $C_{x,y}$ is an interval which contains $q^*_{\mathrm{lo},x,y},q^*_{\mathrm{hi},x,y}$ with high probability.

In practice, we were not able to find a way to compute the credible interval and posterior means exactly.  It seems difficult to even sample the conditional random variables $P,H$.  Common approaches to this type of problem involve Markov-Chain Monte Carlo and Variational Inference, but we were unable to make these approaches work in practice.  It seems nontrivial to work with the condition $E(p,h)=0$ that formalizes the MLM assumption.  We instead take a somewhat na\"ive approach.  We start with $\tilde P,\tilde H$, which carry the distribution of $p^*,h^*$ conditioned on the observed data (but not conditioned on the fact that the MLM assumption holds).  We then simply project $\tilde H$ to the space of parameters which are consistent with the MLM assumption.  That is, we take $\hat H$ to be the solution of the minimization problem $\min_{h'}D(\tilde H|h')$, subject to the constraint that $E(\tilde P,h')=0$.  Optimization details can be found in Appendix \ref{sec:optproblems}.  We hope that the distribution of $\tilde P,\hat H$ will approximate the desired posterior $P,H$.  We then use samples from this distribution to compute what we need.  The implications of this approximation are difficult to reason about; we hope further study will clarify when this is a reasonable procedure.  In the mean time, we content ourselves that it is a practical solution to a difficult problem.

A frequentist point of view might provide a valuable counterpoint to our Bayesian analysis.  However, we were unable to produce effective confidence intervals for the link parameters.  There are at least two difficulties: the link is unidentifiable and the true parameter values may lie on the edge of the parameter space.  These challenges make traditional appproaches based on asymptotic normality difficult to apply.  Perhaps future work could uncover a way to overcome these challenges.

@@

\caption{{\bf When can the MLM accurately estimate the link?} \label{fig:simI}  We use simulations to test the MLM.  In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, sampling procedures, and the link parameters.  In each trial our `ground truth' link parameters picked uniformly at random from the parameter space.  The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct sampling procedures.  In subfigure (b) we consider the MLM's credible intervals.  For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM confidence interval $C_{x,y}$.  We show the interval (in black) centered around the ground truth (in red).  Since the link parameters lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).  The intervals mostly contain the truth, and get narrower with more samples as long as there are enough distinct sampling procedures.   In subfigure (c) we focus on the case of exactly $4$ sampling procedures.  In this case it is only possible to get very accurate estimates if the link takes on a particular form.  If the link is independent (i.e.\ $X,Y$ are independent) or if the link is invertible and deterministic (e.g.\ $X=Y$) then we can determine what the link is.  However, if the truth is more complex, small numbers of sampling procedures may make it impossible to determine the true link; this issue is discussed in detail in Section \ref{sec:casestudies}.}
}

We used simulations to test the two parts of the MLM: the point estimators $\hat q(y|x)$ and credible intervals $C_{x,y}$.  We would ordinarily hope for two things:
\begin{description}
    \item[Estimator convergence] As we obtain more samples, $\hat q(y|x)$ should converge to the true link $q^*(y|x)$ for each $x,y$.
    \item[Conservative coverage] $q^*(y|x) \in C_{x,y}$ with high probability.
\end{description}
Our simulations yielded good news and bad news.  The good news: the MLM generally achieves conservative coverage.  In many cases, the MLM also achieves estimator convergence.  The bad news: in many cases the MLM fails to achieve estimator convergence.  Indeed, we can show that the identifiability issues block any possibility of success in some cases; we detail this problem in Appendix \ref{sec:casestudies}.  The following two simulations give a good overview of the relevant issues:

\begin{itemize}
    \item How does MLM performance depend on the number of samples and the number of sampling procedures?  We suppose we have two measurement procedures which return categorical measurements among six categories.  Thus the first procedure yields $X\in \{1,2,3,4,5,6\}$ and the second yields $Y\in \{1,2,3,4,5,6\}$.  To see how the method performs in different circumstances, we will run many trials.  In each trial we fix the number of sampling procedures and pick `ground truth' parameters uniformly at random from the parameter space.  We then simulate a dataset from these ground truth parameters, fixing the total number of samples (spreading these samples equally among all combinations of sampling procedure and measurement procedure).  Finally, we apply the MLM to the simulated data to get the point estimates $\tilde q(y|x)$ and the credible intervals $C_{x y}$.  We measure the overall estimator convergence using a kind of total variation distance:
    \[
    \mathrm{Error}(\tilde q,q^*)=\frac{1}{12}\sum_{x=1}^{6}\sum_{y=1}^{6} |\tilde q(y|x) - q^*(y|x)|
    \]
    This error ranges between zero and one.  Zero error indicates that $\tilde q=q^*$. An error of one indicates that the estimate has completely incorrect beliefs about the probability mass, i.e. $q^*(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.  We also look at the MLM credible intervals and see how often they cover the true parameters.

    The results are summarized in Figure \ref{fig:simI}.  In trials with more samples, the estimator generally has lower error.  However, to make the error actually converge to zero we need at least six distinct sampling procedures.  This figure also shows that the confidence intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.   With many samples and sampling procedures, the intervals are small and concentrated around the truth.  With fewer samples or sampling procedures, the intervals are sometimes forced to be larger.  However, our uncertainty is not the same for every parameter.  In some cases we are able to obtain a very tight credible interval for one particular parameter even though the overall estimator error is high.

    \item How does estimator convergence depend upon the link parameters themselves?  In each trial of this simulation we use four sampling procedures and our measurement procedures yield one of six categories.  In the previous simulations we saw that estimator convergence was generally impossible in this situation, due to the small number of sampling procedures.  However, in our previous simulations we picked the link parameters uniformly from parameter space.  Now we will be more choosey.   On one extreme, we will produce trials where the link makes $X,Y$ independent, i.e. $q^*(y|x)=q^*(y|x')$ for every $x,x'$.  On the other extreme, we will have trials where $X,Y$ are deterministically related by the equation $X=Y$.  We will also consider every link ``in-between'' these two extremes (found by convex combinations).  Figure \ref{fig:simI} shows that estimator convergence is actually possible in the two extreme cases.  However, estimator convergence fails for the in-between cases.  We leave a complete mathematical understanding of this result to future work.  For the purposes of this paper, we content ourselves by noting one interesting case: let the measurements return one of $2^k$ categories and let us use only $k+1$ carefully chosen sampling procedures.  Now suppose the link defines any invertible deterministic function between $X$ and $Y$.  In this case, with enough enough samples, we can determine both that the relationship is deterministic and the exact specification of the invertible function.  This result is proven in Appendix \ref{sec:casestudies}, Theorem \ref{thm:miracle}.
\end{itemize}

@@

\figgy{
\hfill{}\figgygr{.95}{../images/allenNlxy}\hfill{}
\caption{{\bf The input to the Markov Link Method: one experiment for each measurement procedure and each sampling procedure.}  Here we show a portion of the real-world data to which we applied the MLM.  The Allen Institute gathered neurons from the visual cortex of mice, using a variety of cre/lox-based sampling procedure (cf.\ \cite{tasic2017shared}).  Each strategy was designed to bring out different subpopulations of cells.  Neurons were measured to determine their `type,' using one of two procedures: Standard or Patch.  Standard recognizes 104 different types of neurons.  Patch has a coarser notion of cell-type, distinguishing only 10 types.  For each experiment, we tabulated the number of cells assigned to each type.  Above we show a subset of these results through heatmaps; the color of each square indicates the number of specimens found to have a particular type.  Using this kind of data, our task is to calibrate the two classification protocols.  That is, we want to be able to ask question of the following form: ``if a neuron is classified as being of type `Peri Kcnj8' by Standard, how might it have been classified by Patch?'' \label{fig:alleninput}}
}

@@

Our motivation for this problem arose from the Allen Institute's investigation of neuronal cell-types.  The institute has a variety of methods for examining a neuron and determining what kind of neuron it is.  However, many of these procedures destroy the neurons in the process of measuring them.  We'll look at two procedures in particular, which we'll call `Standard' and 'Patch.'  Standard uses a single-cell RNA sequencing pipeline to determine transcriptomic expression in the cell and determines the cell-type based on these results.  Patch also uses transcriptomic information from RNA sequencing.  Patch additionally obtains electrophysiological and morphological properties of the neuron.  This comes at the cost of a degraded transcriptomic signal, requiring new methods to estimate the cell-type.  We refer the reader to \citep{tasic2017shared} for details on the two methods.

@@

Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: a variety of sampling procedures.  Using a cre/lox system (cf.\ \citep{tasic2017shared}) they were able to pick out various overlapping subpopulations of neurons using different procedures.  Each procedure was expected to yield different proportions of the different cell-types.  For each sampling procedure and each measurement procedure, many specimens were sampled and their cell-types determined.  The result of this process was two tables, parts of which are shown in Figure \ref{fig:alleninput}.  While it seemed clear that these tables should say something about the calibration, it was not obvious how to best use this information.  It was for this purpose that the MLM was developed.

@@

In future work, it would be interesting to use additional assumptions to help us estimate calibrations.  For example, as it stands the Markov Link Method will only yield narrow confidence intervals if each measurement tool returns one of a finite number of measurement outcomes.  In some cases, we may believe that similar measurement values should have similar probabilities.  Such smoothness assumptions would make it possible to apply the MLM to measurement tools which can return a continuum of values.

More generally, it is clear that good assumptions can help us get real insight for tough problems. Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, credible regions can be constructed.  By probing these regions carefully, we can learn what the data actually has to say and what experiments will help us learn more.

@@
